A unified pseudo-$C_\ell$ framework

### 2019-02-13 doc/Paper/main.tex
Updated to gsl-2.5 and added JS acks

\RequirePackage{docswitch}
\setjournal{\flag}

\documentclass[usenatbib]{mnrasb}

% Custom commands from LSST DESC, see texmf/styles/lsstdesc_macros.sty
\usepackage{lsstdesc_macros}
\usepackage{graphicx}

@@

\usepackage{lipsum}
\usepackage{lineno}
\linenumbers

\usepackage{etoolbox}

\newtoggle{DEBUG}
\toggletrue{DEBUG}

\iftoggle{DEBUG}{
\newcommand{\chg}[1]{{\textcolor{blue}{#1}}}
\newcommand{\as}[1]{{\textcolor{red}{[\textbf{AS}: #1]}}}
\newcommand{\dam}[1]{{\textcolor{green}{[\textbf{DA}: #1]}}}
\newcommand{\js}[1]{{\textcolor{orange}{[\textbf{JS}: #1]}}}
\newcommand{\todo}[1]{{\bf TODO: #1}}
}{
\newcommand{\chg}[1]{}
\newcommand{\as}[1]{}
\newcommand{\dam}[1]{}
\newcommand{\js}[1]{}
\newcommand{\todo}[1]{}
}

\begin{document}

\title{A unified pseudo-$C_\ell$ framework}

\maketitlepre

@@

\dockeys{methods: data analysis -- methods: numerical}

\maketitlepost

  \section{Introduction}\label{sec:intro}
    Two-point functions have proven to be the most useful summary statistic for cosmology in terms of data compression, due to the near-Gaussian nature of the modes of the cosmic density field that can be easily analyzed to extract cosmological information~\citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,2001PhRvD..64f3001T}. Since the only true observables in any astronomical dataset are the intensity and polarization as a function of frequency $\nu$ and sky position $\nv$, angular 2-point correlators play a particularly central role in most cosmological analyses, both in the form of configuration-space two-point correlation functions $\xi(\theta)$ and as Fourier-space\footnote{In this paper we use the term ``Fourier'' to describe both the flat space Fourier expansion as well as harmonic-space expansion on the sphere. We will also use the term ``power spectrum'' to describe both the auto power spectrum of a field and cross-power spectra between two fields. } power spectra $C_\ell$ \citep{2012MNRAS.427.1891A,2011PhRvD..84f3505B}.  Although correlation functions are often easier to estimate from complex observations~\citep{1993ApJ...412...64L}, a Fourier space analysis is advantageous for many other reasons. The most important is that Fourier space is a natural space for working with statistics of translation-invariant fields. Therefore, Fourier space analysis has a clearer separation between linear and non-linear scales, as well as Gaussian and non-Gaussian modes~\citep{2008PhRvD..77j3013H}. Moreover, individual band-power measurements are typically significantly less correlated compared to configuration-space analysis which makes covariance matrix estimation easier. This has motivated a large body of literature aimed at designing optimal and efficient power spectrum estimators and used extensively in Cosmic Microwave Background (CMB) two-point measurements.

@@

        In the standard incarnation of the optimal quadratic estimator~\citep{1998PhRvD..57.2117B}, the measurement is produced on inverse covariance weighted data, namely
        \begin{equation}
          {\sf F_{ij}} B_j = \frac{1}{2}\va^\dagger \tilde{\sf C}^{-1}\,{\sf P}_i\,\tilde{\sf C}^{-1} \va -
          \frac{1}{2}\rm{Tr}\left[\tilde{\sf C}^{-1}{\sf P}_i\tilde{\sf C}^{-1}\,{\sf N} \right],
        \end{equation}
        where ${\sf F}$ is the Fisher matrix, the covariance matrix is modeled as a linear sum over band-power parameters $B_i$, i.e.
        \begin{equation}\label{eq:bpws_qmv}
          \tilde{\sf C} = \sum_i {\sf P}_i\,B_i + {\sf N},
        \end{equation}
        (which defines the response matrices ${\sf P}_i$) and the last term in Eq. \ref{eq:bpws_qmv} corresponds to a bias of the convolved estimates due to the presence of noise. Thus, this optimal estimator corresponds to i) inverse-variance weighting the data, ii) calculating the spectrum of the weighted data using the response matrices and iii) deconvolving these estimates using a mode-coupling matrix which in this particular case happens to be the same as Fisher matrix. The main computational challenge when implementing this estimator usually lies in storing the covariance matrix and its inverse (or computing it on the fly), particularly in order to estimate the Fisher matrix. This motivates the pseudo-$C_\ell$ estimator, which replaces these computationally very intensive steps with numerically more efficient ones at the expense of optimality.

@@

      \subsubsection{Bandpowers}\label{sssec:maths.pcl101.bpws}
        Given the loss of information implicit in masking the originally curved-sky field, it is in general not possible to invert the mode-coupling matrix in Eq. \ref{eq:modecoup} directly. One solution to this problem is to convolve the theory prediction with the mode-coupling matrix at the likelihood level. A more usual approach is instead to bin the coupled pseudo-$C_\ell$ into bandpowers. A bandpower $q$ is defined as a set of $N_q$ multipoles $\vec{\ell}_q\equiv(\ell_q^1,...,\ell_q^{N_q})$ and a set of weights defined at those multipoles $w^\ell_q$, and normalized such that $\sum_{\ell\in\vec{\ell}_q}w_q^\ell=1$. The $q$-th bandpower for the coupled pseudo-$C_\ell$ is then defined as
        \begin{align}\label{eq:bpws}
          {\rm vec}\left[\tilde{\sf B}^{ab}_q\right]&\equiv\sum_{\ell\in\vec{\ell}_q}w_q^\ell\,{\rm vec}\left[{\rm PCL}_{\ell}({\bf a}^v_{\bf l},{\bf b}^w_{\bf l})\right]\\\nonumber
          \left\langle{\rm vec}\left[\tilde{\sf B}^{ab}_q\right]\right\rangle&=\sum_{\ell\in\vec{\ell}_q}w_q^\ell\sum_{\ell'}{\sf M}^{s_as_b}_{\ell\ell'}\,{\rm vec}\left[{\sf C}^{ab}_{\ell'}\right],
        \end{align}
        where in the second line we have taken the expectation value of first one. One then proceeds by assuming that the true power spectrum is a step-wise function, taking constant values over the multipoles corresponding to each bandpower: ${\sf C}^{ab}_\ell\equiv\sum_q{\sf B}^{ab}_q\Theta(\ell\in\vec{\ell}_q)$ (where $\Theta$ is a binary step function). Inserting this in the equation above, it is then possible to find an unbiased estimator for ${\sf B}^{ab}_q$ in terms of the binned pseudo-bandpowers $\tilde{\sf B}^{ab}_q$:
        \begin{equation}\label{eq:decoup}
          {\rm vec}\left[{\sf B}^{ab}_q\right]=\sum_{q'}\left({\cal M}^{s_as_b}\right)^{-1}_{qq'}\,{\rm vec}\left[\tilde{\sf B}^{ab}_{q'}\right],
        \end{equation}
        where the binned coupling matrix ${\cal M}^{s_as_b}$ is
        \begin{equation}
          {\cal M}^{s_as_b}_{qq'}=\sum_{\ell\in\vec{\ell}_q}\sum_{\ell'\in\vec{\ell}_{q'}}w_q^\ell\,{\sf M}^{s_as_b}_{\ell\ell'}.
        \end{equation}

@@

    \subsection{Contaminant deprojection}\label{ssec:maths.deproj}
      In a typical data-analysis case, we deal with contaminants that pollute the measured signal. Usually, we know something about these contaminants. As an example, in case of CMB analysis, we might have a map of a given foreground at a different frequency which can be marginalized out. But even when we do not, we often have a tracer map of a given effect, for example, the map of dust might tell us something about systematic errors related to imperfect correction to dust reddening, or a map of co-added PSF fluctuations could be used to remove the possible correlations between measured ellipticities. Even if the dust map is non-linearly and perhaps stochastically related to the amount of reddening, as long as that relation is local in space, we get a linear contamination on large scales, very much like complex astrophysics of galaxy formation nevertheless reduces to linear biasing on very large scales (see Appendix \ref{app:lincont}). Thus, we generically deal with a linear model of the kind:
      \begin{equation}\label{eq:aobs}
        {\bf a}^{\rm obs}(\nv)={\bf a}^v+\sum_{i=1}^{N_a}\alpha_i{\bf f}^i(\nv),
      \end{equation}
      where the observed map ${\bf a}^{\rm obs}$ is composed of a masked signal map that receives contributions from $N_a$ known contaminants ${\bf f}^i$ and where the contamination coefficients $\alpha_i$ are unknown (note that we have implicitly applied the same mask to the contaminant templates ${\bf f}^i$). The deprojection of these contaminants from the data can be done and propagated into the estimated power spectrum in a natural way in the case of quadratic estimators \citep{1992ApJ...398..169R,2004PhRvD..69l3003S,2016MNRAS.456.2095E}, and the method can be extended to the pseudo-$C_\ell$ algorithm \citep{2017MNRAS.465.1847E}. The process is illustrated in Figure \ref{fig:lss_cont}. A best-fit value for the coefficients $\alpha_i$ can be found assuming \emph{uncorrelated weights} across the map\footnote{The assumption of uncorrelated weights is the key difference between pseudo-$C_\ell$ and optimal quadratic estimators \citep{2013MNRAS.435.1857L}. Note that, since the templates are already multiplied by the mask, the weights are not equal across the map, but correspond to the mask values in general.}
      \begin{align}\label{eq:deproj_coeff}
        &\hat{\alpha}_i=F_{ij}\int d\nv\,{\bf f}^{j\dag}(\nv)\,{\bf a}^{\rm obs}(\nv),\\\nonumber
        &({\sf F}^{-1})_{ij}\equiv\int d\nv\, {\bf f}^{i\dag}(\nv)\,{\bf f}^j(\nv),
      \end{align}
      where there is an implicit summation over the repeated index $j$ in Eq. \ref{eq:deproj_coeff} (which we will omit from now on). The clean map is then given by
      \begin{align}
        {\bf a}^c(\nv)
        &\equiv{\bf a}^{\rm obs}(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^{\rm obs}(\nv')
      \end{align}
      which, when substituting Eq. \ref{eq:aobs}, yields
      \begin{align}\label{eq:clean}
        {\bf a}^c(\nv)&={\bf a}^v(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^v(\nv').
      \end{align}      
      The harmonic coefficients of ${\bf a}^c$ are therefore given by
      \begin{equation}\label{eq:clean_f}
        {\bf a}^c_{\bf l}={\bf a}^v_{\bf l}-{\bf f}^i_{\bf l}\,F_{ij}\sum_{{\bf l}'}{\bf f}^{j\dag}_{{\bf l}'}{\bf a}^v_{{\bf l}'}
      \end{equation}

@@

    \subsection{\texorpdfstring{$E/B$}{E/B} purification}\label{ssec:maths.pureb}
      As described in \cite{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2010PhRvD..82b3001Z,2010A&A...519A.104K,2011PhRvD..83h3003B}, $B$-mode purification refers to the map-level removal of the contamination from $E$-modes in a the $B$-mode component of a given map caused by an incomplete sky coverage and vice-versa\footnote{Note that this is a different effect from the leakage between different polarized and unpolarized channels in CMB observations due to instrumental effects (asymmetric beams, pair differencing, polarization angle calibration etc.)}. The procedure is particularly useful in situations in which the $E$-mode component of the signal is significantly larger than the $B$ modes, as is the case in CMB. In this case, removing the leakage at the power spectrum level (i.e. the standard pseudo-$C_\ell$ approach) produces a sub-optimal estimator in which the variance in the $B$-mode power spectrum is dominated by the variance of the leaked $E$-modes. This section presents the purification method and the associated modifications to the pseudo-$C_\ell$ algorithm as described in the references above. In this section we describe the algorithm in terms of purifying $B$-modes from $E$-mode contamination, but we note that the reverse, i.e. purifying $E$-modes from $B$-mode contamination is conceptually exactly the same. Our discussion is also specific for spin-2 fields.

@@

@@

        \subsubsection{Beam deconvolution}\label{sssec:maths.beam_noise.beam}
          Including the effect of an spherically symmetric instrumental beam amounts to a simple redefinition of the mode-coupling matrix
          \begin{equation}
            {\sf M}^{s_as_b}_{\ell_1\ell_2}\rightarrow{\sf M}^{s_as_b}_{\ell_1\ell_2}W_{\ell_2}^{ab},
          \end{equation}
          where $W^{ab}_\ell$ is the product of the harmonic transform of the beams for maps $a$ and $b$. Note that since any continuous tangent vector field on the sphere must, at some location, be zero (the so-called hairy ball theorem), it is impossible to have an anisotropic beam that is uniform over the entire sphere.  However, it is possible that the beam is both anisotropic and spatially-varying. In this case, the pseudo-$C_\ell$ power spetrum estimation will produce biased estimates that will be suppressed by an \emph{effective} beam. Correcting for the beam in such cases can also be done analytically, but we leave this for future work.  In practice, if required, the
          suppression factor can be found by applying the PCL estimator to a set of mock input maps with the same underlying signal but  with and without application of the beam smoothing (to both data maps and contaminant maps) and taking the ratio.

@@

      \subsubsection{Galaxy clustering and weak lensing} \label{sssec:validation.suite.lss}
        \begin{figure}
          \centering
          \includegraphics[width=0.99\columnwidth]{cls_lss}
          \caption{Signal and noise power spectra used in the LSS validation set. Dark blue, cyan and red curves show the spectra associated with the $\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$ correlations respectively, where $\gamma_E$ is the lensing $E$ mode. All other cross-correlations are zero. The dashed lines show the noise power spectra, associated with shot noise (red) and intrinsic shape scatter (blue).} \label{fig:lss_cls}
        \end{figure}
        \begin{figure*}
          \centering
          \includegraphics[width=0.9\textwidth]{maps_lss_sph}
          \caption{Inputs for the large-scale structure, curved-sky validation set. From top to bottom and left to right: sky mask (1), Gaussian realization of the galaxy overdensity $\delta$ (2) and cosmic shear $\gamma_i$ (3,4), dust contaminant for $\delta$ (5), star contaminant for $\delta$ (6), PSF contaminant for $\gamma$ (7, $Q$ component) and small-scale contaminant for $\gamma$ (8, $Q$ component).}
          \label{fig:lss_mocks_sph}
        \end{figure*}
        \begin{figure*}
          \centering
          \includegraphics[width=0.9\textwidth]{maps_lss_flat}
          \caption{Same as Fig. \ref{fig:lss_mocks_sph} for the large-scale structure, flat-sky validation set.}
          \label{fig:lss_mocks_flat}
        \end{figure*}
        \begin{figure}
          \centering
          \includegraphics[width=0.99\columnwidth]{cls_cont_lss}
          \caption{{\sl Top:} signal power spectrum (solid dark blue) for the auto-correlation of the galaxy overdensity, as well as the contaminants associated to dust (dashed cyan) and stars (dot-dashed red). {\sl Middle:} $E$-mode lensing power spectrum (solid dark blue) and the spectra of the PSF and small-scale contaminants used in our validation set (dashed cyan and dot-dashed red respectively). {\sl Bottom:} same as middle panel for the $B$-mode power spectrum. Note that in this case the input signal power spectrum is $0$. In all cases the orange dotted line shows the combination of signal and contaminants.} \label{fig:lss_cont_cls}
        \end{figure}
        The first validation suite exemplifies the use of pseudo-$C_\ell$ algorithms in the analysis of large-scale structure data. Current and next-generation imaging surveys such as DES~\citep{2016MNRAS.460.1270D},
        KiDS~\citep{2017A&A...604A.134D}, HSC~\citep{2018PASJ...70S...4A}, LSST~\citep{2008arXiv0805.2366I} and Euclid~\citep{2011arXiv1110.3193L} will constrain the growth and geometry of structure through cosmic time by using the overdensity of galaxies as a proxy for the matter inhomogeneities (a spin-0 field) and their correlated shape distortions (a spin-2 field at first order), mostly caused by weak lensing. These measurements will mostly follow a tomographic approach, using the angular auto- and cross-power spectra between these observables across different bins of photometric redshift \citep{2017arXiv170609359K,2018MNRAS.474.4894J,2017arXiv170801530D}.

@@

        For this validation suite we therefore use a mock dataset mimicking a typical observation of an LSST-like survey in a particular redshift bin. Slightly different set-ups are used for the curved-sky and flat-sky validation suites. The characteristics of these data are as follows:

@@

          \item {\bf Noise:} for the curved-sky case we generate inhomogeneous noise realizations. To do this, we first generate maps with Gaussian, white and homogeneous noise with a pixel variance given by $1/n_{\rm pix}$ for $\delta$ and $\sigma_\gamma^2/n_{\rm pix}$ for each shear component, where $n_{\rm pix}$ is the average number of sources in a pixel's area, and $\sigma_\gamma=0.28$ is then intrinsic ellipticity scatter per component. These noise realizations are then locally rescaled by a factor $f(\nv)=f_{\rm star}(\nv)\,f_{\rm depth}(\nv)$ made up of two components:

@@

  \input{contributions}

  DA acknowledges support from the Beecroft trust and from the Science and Technology Facilities Council (STFC) through an Ernest Rutherford Fellowship, grant reference ST/P004474/1.
  \input{desc-tex/ack/standard}

\bibliographystyle{mnrasb}
\bibliography{main}

>>>>

\documentclass[usenatbib]{mnrasb}
\bibliographystyle{mnrasb}
\pdfoutput=1
\pdfminorversion=5

%\documentclass[usenatbib]{mnrasb}

% Custom commands from LSST DESC, see texmf/styles/lsstdesc_macros.sty
%\usepackage{lsstdesc_macros}
\usepackage{graphicx}

@@

\title[A unified pseudo-$C_\ell$ framework]{A unified pseudo-$C_\ell$ framework}
  \author[D. Alonso et al.]{David Alonso$^{1,2}$\thanks{david.alonso@physics.ox.ac.uk}, Javier Sanchez$^{3}$, An\v{z}e Slosar$^{4}$ \newauthor{(The LSST Dark Energy Science Collaboration)}\\
  $^{1}$ School of Physics and Astronomy, Cardiff University, The Parade, Cardiff, CF24 3AA, United Kingdom\\
  $^{2}$ Department of Physics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, United Kingdom\\
  $^{3}$ Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA\\
  $^{4}$ Brookhaven National Laboratory, Physics Department, Upton, NY 11973, USA}

\begin{document}
  \date{\today}
  \pagerange{1--27} \pubyear{2018}
  \maketitle

@@

  \begin{keywords}
    methods: data analysis -- methods: numerical
  \end{keywords}

  \section{Introduction}\label{sec:intro}
    Two-point functions have proven to be the most useful summary statistic for cosmology in terms of data compression, due to the near-Gaussian nature of the modes of the cosmic density field that can be easily analyzed to extract cosmological information~\citep{1994ApJ...430L..85G,1995PhRvL..74.4369B,1997PhRvD..55.5895T,2001PhRvD..64f3001T}. Since the only true observables in any astronomical dataset are the intensity and polarization as a function of frequency $\nu$ and sky position $\nv$, angular 2-point correlators play a particularly central role in most cosmological analyses, both in the form of configuration-space two-point correlation functions $\xi(\theta)$ and as Fourier-space\footnote{In this paper we use the term ``Fourier'' to describe both the flat space Fourier expansion as well as harmonic-space expansion on the sphere. We will also use the term ``power spectrum'' to describe both the auto power spectrum of a field and cross-power spectra between two fields. } power spectra $C_\ell$ \citep{2012MNRAS.427.1891A,2011PhRvD..84f3505B}.  Although correlation functions are often easier to estimate from complex observations~\citep{1993ApJ...412...64L}, a Fourier space analysis is advantageous for many other reasons. The most important is that Fourier space is a natural space for working with statistics of translation-invariant fields. Therefore, Fourier space analysis has a clearer separation between linear and non-linear scales, as well as Gaussian and non-Gaussian modes~\citep{2008PhRvD..77j3013H}. Moreover, individual band-power measurements are typically significantly less correlated compared to configuration-space analysis which makes covariance matrix estimation easier. This has motivated a large body of literature aimed at designing optimal and efficient power spectrum estimators and used extensively in Cosmic Microwave Background (CMB) two-point measurements.

@@

        In the standard incarnation of the optimal quadratic estimator~\citep{1998PhRvD..57.2117B}, the measurement is produced on inverse covariance weighted data, namely
        \begin{equation}\label{eq:fisher_bpw}
          {\sf F_{ij}} B_j = \frac{1}{2}\va^\dagger \tilde{\sf C}^{-1}\,{\sf P}_i\,\tilde{\sf C}^{-1} \va -
          \frac{1}{2}\rm{Tr}\left[\tilde{\sf C}^{-1}{\sf P}_i\tilde{\sf C}^{-1}\,{\sf N} \right],
        \end{equation}
        where ${\sf F}$ is the Fisher matrix, the covariance matrix is modeled as a linear sum over band-power parameters $B_i$, i.e.
        \begin{equation}\label{eq:bpws_qmv}
          \tilde{\sf C} = \sum_i {\sf P}_i\,B_i + {\sf N},
        \end{equation}
        (which defines the response matrices ${\sf P}_i$) and the last term in Eq. \ref{eq:fisher_bpw} corresponds to a bias of the convolved estimates due to the presence of noise. Thus, this optimal estimator corresponds to i) inverse-variance weighting the data, ii) calculating the spectrum of the weighted data using the response matrices and iii) deconvolving these estimates using a mode-coupling matrix which in this particular case happens to be the same as Fisher matrix. The main computational challenge when implementing this estimator usually lies in storing the covariance matrix and its inverse (or computing it on the fly), particularly in order to estimate the Fisher matrix. This motivates the pseudo-$C_\ell$ estimator, which replaces these computationally very intensive steps with numerically more efficient ones at the expense of optimality.

@@

      \subsubsection{Bandpowers}\label{sssec:maths.pcl101.bpws}
        Given the loss of information implicit in masking the originally curved-sky field, it is in general not possible to invert the mode-coupling matrix in Eq. \ref{eq:modecoup} directly. One solution to this problem is to convolve the theory prediction with the mode-coupling matrix at the likelihood level. A more usual approach is instead to bin the coupled pseudo-$C_\ell$ into bandpowers. A bandpower $q$ is defined as a set of $N_q$ multipoles $\vec{\ell}_q\equiv(\ell_q^1,...,\ell_q^{N_q})$ and a set of weights defined at those multipoles $w^\ell_q$, and normalized such that $\sum_{\ell\in\vec{\ell}_q}w_q^\ell=1$. The $q$-th bandpower for the coupled pseudo-$C_\ell$ is then defined as
        \begin{align}\label{eq:bpws}
          {\rm vec}\left[\tilde{\sf B}^{ab}_q\right]&\equiv\sum_{\ell\in\vec{\ell}_q}w_q^\ell\,{\rm vec}\left[{\rm PCL}_{\ell}({\bf a}^v_{\bf l},{\bf b}^w_{\bf l})\right]\\\nonumber
          \left\langle{\rm vec}\left[\tilde{\sf B}^{ab}_q\right]\right\rangle&=\sum_{\ell\in\vec{\ell}_q}w_q^\ell\sum_{\ell'}{\sf M}^{s_as_b}_{\ell\ell'}\,{\rm vec}\left[{\sf C}^{ab}_{\ell'}\right],
        \end{align}
        where in the second line we have taken the expectation value of the first one. One then proceeds by assuming that the true power spectrum is a step-wise function, taking constant values over the multipoles corresponding to each bandpower: ${\sf C}^{ab}_\ell\equiv\sum_q{\sf B}^{ab}_q\Theta(\ell\in\vec{\ell}_q)$ (where $\Theta$ is a binary step function). Inserting this in the equation above, it is then possible to find an unbiased estimator for ${\sf B}^{ab}_q$ in terms of the binned pseudo-bandpowers $\tilde{\sf B}^{ab}_q$:
        \begin{equation}\label{eq:decoup}
          {\rm vec}\left[{\sf B}^{ab}_q\right]=\sum_{q'}\left({\cal M}^{s_as_b}\right)^{-1}_{qq'}\,{\rm vec}\left[\tilde{\sf B}^{ab}_{q'}\right],
        \end{equation}
        where the binned coupling matrix ${\cal M}^{s_as_b}$ is
        \begin{equation}
          {\cal M}^{s_as_b}_{qq'} \equiv \sum_{\ell\in\vec{\ell}_q}\sum_{\ell'\in\vec{\ell}_{q'}}w_q^\ell\,{\sf M}^{s_as_b}_{\ell\ell'}.
        \end{equation}

@@

    \subsection{Contaminant deprojection}\label{ssec:maths.deproj}
      In a typical data-analysis case, we deal with contaminants that pollute the measured signal. Usually, we know something about these contaminants. As an example, in case of CMB analysis, we might have a map of a given foreground at a different frequency which can be marginalized out\footnote{It is worth noting that the formalism described here assumes that a given contaminant template has no significant contribution from the signal itself. This might not be the case, for example, if one blindly took a high-frequency sky map to serve as a template of dust, since that may still contain a significant CMB contribution. Care must therefore be exercised when defining the contaminant templates to marginalize over.}. But even when we do not, we often have a tracer map of a given effect, for example, the map of dust might tell us something about systematic errors related to imperfect correction to dust reddening, or a map of co-added point spread function (PSF) fluctuations could be used to remove the possible correlations between measured ellipticities. Even if the dust map is non-linearly and perhaps stochastically related to the amount of reddening, as long as that relation is local in space, we get a linear contamination on large scales, very much like complex astrophysics of galaxy formation nevertheless reduces to linear biasing on very large scales (see Appendix \ref{app:lincont}). Thus, we generically deal with a linear model of the kind:
      \begin{equation}\label{eq:aobs}
        {\bf a}^{\rm obs}(\nv)={\bf a}^v+\sum_{i=1}^{N_a}\alpha_i{\bf f}^i(\nv),
      \end{equation}
      where the observed map ${\bf a}^{\rm obs}$ is composed of a masked signal map that receives contributions from $N_a$ known contaminants ${\bf f}^i$ and where the contamination coefficients $\alpha_i$ are unknown (note that we have implicitly applied the same mask to the contaminant templates ${\bf f}^i$). The deprojection of these contaminants from the data can be done and propagated into the estimated power spectrum in a natural way in the case of quadratic estimators \citep{1992ApJ...398..169R,2004PhRvD..69l3003S,2016MNRAS.456.2095E}, and the method can be extended to the pseudo-$C_\ell$ algorithm \citep{2017MNRAS.465.1847E}. The process is illustrated in Figure \ref{fig:lss_cont}. A best-fit value for the coefficients $\alpha_i$ can be found assuming \emph{uncorrelated weights} across the map\footnote{The assumption of uncorrelated weights is the key difference between pseudo-$C_\ell$ and optimal quadratic estimators \citep{2013MNRAS.435.1857L}. Note that, since the templates are already multiplied by the mask, the weights are not equal across the map, but correspond to the mask values in general.}
      \begin{align}\label{eq:deproj_coeff}
        &\hat{\alpha}_i=F_{ij}\int d\nv\,{\bf f}^{j\dag}(\nv)\,{\bf a}^{\rm obs}(\nv),\\\nonumber
        &({\sf F}^{-1})_{ij}\equiv\int d\nv\, {\bf f}^{i\dag}(\nv)\,{\bf f}^j(\nv),
      \end{align}
      where there is an implicit summation over the repeated index $j$ in Eq. \ref{eq:deproj_coeff} (which we will omit from now on). The clean map is then given by
      \begin{align}
        {\bf a}^c(\nv)
        &\equiv{\bf a}^{\rm obs}(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^{\rm obs}(\nv')
      \end{align}
      which, when substituting Eq. \ref{eq:aobs}, yields
      \begin{align}\label{eq:clean}
        {\bf a}^c(\nv)&={\bf a}^v(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^v(\nv').
      \end{align}      
      The harmonic coefficients of ${\bf a}^c$ are therefore given by
      \begin{equation}\label{eq:clean_f}
        {\bf a}^c_{\bf l}={\bf a}^v_{\bf l}-{\bf f}^i_{\bf l}\,F_{ij}\sum_{{\bf l}'}{\bf f}^{j\dag}_{{\bf l}'}{\bf a}^v_{{\bf l}'}
      \end{equation}

@@

    \subsection{\texorpdfstring{$E/B$}{E/B} purification}\label{ssec:maths.pureb}
      As described in \cite{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2010PhRvD..82b3001Z,2010A&A...519A.104K,2011PhRvD..83h3003B}, $B$-mode purification refers to the map-level removal of the contamination from $E$-modes in the $B$-mode component of a given map caused by an incomplete sky coverage and vice-versa\footnote{Note that this is a different effect from the leakage between different polarized and unpolarized channels in CMB observations due to instrumental effects (asymmetric beams, pair differencing, polarization angle calibration etc.)}. The procedure is particularly useful in situations in which the $E$-mode component of the signal is significantly larger than the $B$ modes, as is the case in for the CMB. In this case, removing the leakage at the power spectrum level (i.e. the standard pseudo-$C_\ell$ approach) produces a sub-optimal estimator in which the variance in the $B$-mode power spectrum is dominated by the variance of the leaked $E$-modes. This section presents the purification method and the associated modifications to the pseudo-$C_\ell$ algorithm as described in the references above. In this section we describe the algorithm in terms of purifying $B$-modes from $E$-mode contamination, but we note that the reverse, i.e. purifying $E$-modes from $B$-mode contamination is conceptually exactly the same. Our discussion is also specific for spin-2 fields.

@@

      It is worth noting that, as described in \cite{2009PhRvD..79l3515G}, the performance of $E/B$ purification can be optimized by an appropriate choice of window function. The corresponding optimal spin-0, spin-1 and spin-2 window functions can be estimated using a preconditioned conjugate-gradient method. The resulting spin window functions are no longer related by simple covariant derivatives. The performance of this method, compared with simple apodization methods, will depend on the noise properties of the experiment and the footprint geometry. This optimal method is currently not available in \nmt, and will be implemented in future releases of the code.

@@

        \subsubsection{Beam deconvolution}\label{sssec:maths.beam_noise.beam}
          Including the effect of a spherically symmetric instrumental beam amounts to a simple redefinition of the mode-coupling matrix
          \begin{equation}
            {\sf M}^{s_as_b}_{\ell_1\ell_2}\rightarrow{\sf M}^{s_as_b}_{\ell_1\ell_2}W_{\ell_2}^{ab},
          \end{equation}
          where $W^{ab}_\ell$ is the product of the harmonic transform of the beams for maps $a$ and $b$. Note that since any continuous tangent vector field on the sphere must, at some location, be zero (the so-called hairy ball theorem), it is impossible to have an anisotropic beam that is uniform over the entire sphere.  However, it is possible that the beam is both anisotropic and spatially-varying. In this case, the pseudo-$C_\ell$ power spetrum estimation will produce biased estimates that will be suppressed by an \emph{effective} beam. Correcting for the beam in such cases can also be done analytically, but we leave this for future work.  In practice, if required, the
          suppression factor can be found by applying the PCL estimator to a set of mock input maps with the same underlying signal but  with and without application of the beam smoothing (to both data maps and contaminant maps) and taking the ratio.

@@

      \subsubsection{Galaxy clustering and weak lensing} \label{sssec:validation.suite.lss}
        \begin{figure}
          \centering
          \includegraphics[width=0.99\columnwidth]{cls_lss}
          \caption{Signal and noise power spectra used in the LSS validation set. Dark blue, cyan and red curves show the spectra associated with the $\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$ correlations respectively, where $\gamma_E$ is the lensing $E$ mode. All other cross-correlations are zero. The dashed lines show the noise power spectra, associated with shot noise (red) and intrinsic shape scatter (blue).} \label{fig:lss_cls}
        \end{figure}
        \begin{figure*}
          \centering
          \includegraphics[width=0.99\textwidth]{maps_lss_sph}
          \caption{Inputs for the large-scale structure, curved-sky validation set. From top to bottom and left to right: sky mask (1), Gaussian realization of the galaxy overdensity $\delta$ (2) and cosmic shear $\gamma_i$ (3,4), dust contaminant for $\delta$ (5), star contaminant for $\delta$ (6), PSF contaminant for $\gamma$ (7, $Q$ component) and small-scale contaminant for $\gamma$ (8, $Q$ component).}
          \label{fig:lss_mocks_sph}
        \end{figure*}
        \begin{figure*}
          \centering
          \includegraphics[width=0.9\textwidth]{maps_lss_flat}
          \caption{Same as Fig. \ref{fig:lss_mocks_sph} for the large-scale structure, flat-sky validation set.}
          \label{fig:lss_mocks_flat}
        \end{figure*}
        \begin{figure}
          \centering
          \includegraphics[width=0.99\columnwidth]{cls_cont_lss}
          \caption{{\sl Top:} signal power spectrum (solid dark blue) for the auto-correlation of the galaxy overdensity, as well as the contaminants associated to dust (dashed cyan) and stars (dot-dashed red). {\sl Middle:} $E$-mode lensing power spectrum (solid dark blue) and the spectra of the PSF and small-scale contaminants used in our validation set (dashed cyan and dot-dashed red respectively). {\sl Bottom:} same as middle panel for the $B$-mode power spectrum. Note that in this case the input signal power spectrum is $0$. In all cases the orange dotted line shows the combination of signal and contaminants.} \label{fig:lss_cont_cls}
        \end{figure}
        The first validation suite exemplifies the use of pseudo-$C_\ell$ algorithms in the analysis of large-scale structure data. Current and next-generation imaging surveys such as DES~\citep{2016MNRAS.460.1270D},
        KiDS~\citep{2017A&A...604A.134D}, HSC~\citep{2018PASJ...70S...4A}, LSST~\citep{2008arXiv0805.2366I} and Euclid~\citep{2011arXiv1110.3193L} will constrain the growth and geometry of structure through cosmic time by using the overdensity of galaxies as a proxy for the matter inhomogeneities (a spin-0 field) and their correlated shape distortions (a spin-2 field at first order), mostly caused by weak lensing. These measurements will mostly follow a tomographic approach, using the angular auto- and cross-power spectra between these observables across different bins of photometric redshift \citep{2017arXiv170609359K,2018MNRAS.474.4894J,2017arXiv170801530D}.

@@

        For this validation suite we therefore use a mock dataset mimicking a typical observation of a LSST-like survey in a particular redshift bin. Slightly different set-ups are used for the curved-sky and flat-sky validation suites. The characteristics of these data are as follows:

@@

          \item {\bf Noise:} for the curved-sky case we generate inhomogeneous noise realizations. To do this, we first generate maps with Gaussian, white and homogeneous noise with a pixel variance given by $1/n_{\rm pix}$ for $\delta$ and $\sigma_\gamma^2/n_{\rm pix}$ for each shear component, where $n_{\rm pix}$ is the average number of sources in a pixel's area, and $\sigma_\gamma=0.28$ is the intrinsic ellipticity scatter per component. These noise realizations are then locally rescaled by a factor $f(\nv)=f_{\rm star}(\nv)\,f_{\rm depth}(\nv)$ made up of two components:

@@

  Author contributions are listed below. \\
  David Alonso: Co-led project; wrote the code; contributed to validation. \\
  Javier Sanchez: Co-led project; wrote code; contributed to validation. \\
  An\v{z}e Slosar: Co-led project; contributed to validation. \\
